---
title: "Results"
---

### Code to Process Data and Perform One-Sample t-Test

```{r, eval = FALSE}
food <- read.csv("FAOSTAT_data_4-29-2018.csv")
attach(food)

food_continent <- read.csv("FAOSTAT_data_4-30-2018_continent.csv")
attach(food_continent)

# Initialize 3D array to store matrices of 5-year food supplies for each country 
foodsupply <- array(NA,c(5,length(unique(Country)),length(unique(Year))/5))
# Get dimensions of "foodsupply" array to use later
dims <- dim(foodsupply)

# Initalize matrices to store averages and SDs of foodsupply for all countries included in a 5-year period
country_avgs <- matrix(NA,nrow = length(unique(Year))/dims[1],ncol = length(unique(Country)))
country_sds <- matrix(NA,nrow = length(unique(Year))/dims[1],ncol = length(unique(Country)))

# Initialize matrices to store 5-year continental averages and SDs
continental_avgs <- matrix(nrow = length(unique(Year))/dims[1],ncol = 1)
continental_sds <- matrix(nrow = length(unique(Year))/dims[1],ncol = 1)

# Initialize matrix to store indices of "Year" containing years of current 5-year period 
# (each row is a "Value" from a country having data from specfic year, so up to 44 rows; each column is a year, so 5 columns)
year_inds <- array(NA,c(length(unique(Country)),dims[1],length(unique(Year))/dims[1]))

# Initalize vector to store p-values, and matrix to store confidence intervals from t-tests
p_values <- matrix(nrow = length(unique(Year))/dims[1], ncol = 1)
conf_ints <- matrix(nrow = length(unique(Year))/dims[1], ncol = 2)


for (yy in seq(from = 1,to = length(unique(Year)),by = dims[1])) {
    # Get indices for each year of current 5-year period from "Year" variable
    n <- 0
    for (gg in 1:dims[1]) {
    inds = which(Year == Year[yy + n])
    year_inds[1:length(inds),gg,(yy+4)/dims[1]] <- inds
    n <- n + 1
    }
    
    # Get the food supply(kcal/capita/day) for each country, each year, having data for current 5-year period
    for (ii in 1:dims[1]) {
      foodsupply[ii,,(yy+4)/dims[1]] <- Value[year_inds[,ii,(yy+4)/dims[1]]]
    }
    
    # Calculate average and SD of food suppy for each country over current 5-year period
    for (jj in 1:length(unique(Country))) {
      country_avgs[(yy+4)/dims[1],jj] <- mean(foodsupply[,jj,(yy+4)/dims[1]],na.rm = T)
      country_sds[(yy+4)/dims[1],jj] <- sd(foodsupply[,jj,(yy+4)/dims[1]],na.rm = T)
    }

    # Calculate the average European continental food supply and SD for current 5-year period
    continental_avgs[(yy+4)/dims[1]] <- mean(Value_con[yy:(yy+4)])
    continental_sds[(yy+4)/dims[1]] <- sd(Value_con[yy:(yy+4)])
  
    # Perform one-sample t-test on average food supply/country over current 5-year period
    t = t.test(country_avgs[(yy+4)/dims[1],],mu = continental_avgs[(yy+4)/dims[1]])
    p_values[(yy+4)/dims[1],] <- t$p.value
    conf_ints[(yy+4)/dims[1],] <- t$conf.int
    
}    

```



```{r}
conf_ints_singlecell <- transform(conf_ints,newcol = paste(conf_ints[,1],conf_ints[,2],sep = "-"))
results <- data.frame(p_values,conf_ints_singlecell[,3],null_mean,estimate_mean,
                      row.names = c("1970-1974","1975-1979","1980-1984","1985-1989","1990-1994","1995-1999",
                                    "2000-2004","2005-2009"))
colnames(results) <- c("P-Value","Confidence Interval","Continetnal Mean","True Sample Mean")
kable(results,format = "html",digits = 4,align = 'c',caption = "Results of one-sample t-test on five-year averaged food supplies available per capita, per day in European countries") %>%
   kable_styling(bootstrap_options = "bordered",font_size = 12)

```

The P-values shown in the table above indicate that for a majority (6 out of 8) of the samples, the five-year averaged food supply available each day to every member of a European country's population did not differ significantly from the five-year continetnal mean. However, in two of the samples comprising years 1990-1994 and 2005-2009, the t-tests resulted in P-values less than alpha(0.05), leading to a rejection of the null hypothesis. However, the evidence against the null for the 1990-1994 sample is much stronger than for 2005-2009 because its P-Value is quite a bit smaller.


